---
title: "MSstatsPPI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MSstatsPPI: Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

```{r setup}
library(MSstatsPPI)
```

# Purpose of MSstatsPPI
  
The `MSstatsPPI` package is a member of the `MSstats` family of packages.
It contains a set of functions for interpretation of mass spectrometry (MS) 
statistical analysis results in the context of protein-protein interaction 
(PPI) networks. The package is designed to be used in conjunction with the 
`MSstats` package.

The package provides a function `getSubnetworkFromIndra` that retrieves a
subnetwork of PPIs from the INDRA database based on differential abundance
analysis results.  The function `visualizeSubnetwork` then takes the output
of `getSubnetworkFromIndra` and visualizes the subnetwork.

```{r proteinNetworkDiscovery, eval=FALSE}
input <- data.table::fread(
    system.file("processed_data/groupComparisonModel.csv",
        package = "MSstatsPPI"
    )
)
subnetwork <- getSubnetworkFromIndra(input)
visualizeSubnetwork(subnetwork$nodes, subnetwork$edges)
```
